\documentclass[11pt]{article}
\usepackage{amsmath}
\usepackage{amssymb}
\usepackage[pdftex]{graphicx}
\usepackage{moreverb}


\topmargin -2.0cm
\oddsidemargin 0in
\evensidemargin 0in
%\footheight 0.0pt
\headheight 2\baselineskip
\textwidth 6.2in
\textheight 9.0in


\pagestyle{empty}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{document}
\section*{Benjamin S. Kirk}
\subsection*{List of Graduate Courses}
\begin{itemize}
  \item \textbf{Fluids Courses}
    \begin{enumerate}
    \item ASE 382Q-1 -- \emph{Foundations of Fluid Mechanics}, Fall 2000. \textbf{A}. 
    \item ASE 346 -- \emph{Viscous Fluid Flow}, Fall 2001.  \textbf{A}.
    \item ASE 362K -- \emph{Compressible Fluid Mechanics}, Spring 2001. \textbf{A}.
    \item ASE 382R-3 -- \emph{Hypersonic Aerodynamics}, Fall 2001. \textbf{A}.
    \item ASE 382R-6 -- \emph{Molecular Gas Dynamics}, Spring 2002. \textbf{A}.
    \item ASE 382Q-7 -- \emph{Advanced Problems in Compressible Flow}, Fall 2002. \textbf{A}.
    \end{enumerate}
  \item \textbf{Numerical Methods}
    \begin{enumerate}
    \item ASE 384P-4 -- \emph{Finite Element Methods}, Fall 2000. \textbf{A}.
    \item EM 393N -- \emph{Numerical Methods for Flow \& Transport Problems}, Fall 2000. \textbf{A}.
    \item EM 397 -- \emph{Advanced Methods in Computational Mechanics}, Spring 2001. \textbf{A}.
    \item EM 397 -- \emph{Grid Generation \& Adaptive Grids}, Fall 2001. \textbf{A}.
    \item CAM 383C -- \emph{Numerical Analysis: Numerical Linear Algebra}, Fall 2001. \textbf{B}.
    \item EM 397 -- \emph{Numerical Simulation of Transport in Semiconductors}, Summer 2002. \textbf{A}.
    \item ASE 382Q-8 -- \emph{Lagrangian Methods in Computational Fluid Dynamics}, Fall 2002. \textbf{A}.
    \item EM 397 -- \emph{Advanced Computational Flows \& Transport}, Fall 2002. \textbf{A}.
    \end{enumerate}
  \item \textbf{Other}
    \begin{enumerate}
    \item M 340L -- \emph{Matrices and Matrix Calculations}, Summer 2000. \textbf{A}.
    \item EM 386K -- \emph{Mathematical Methods in Applied Mechanics}, Fall 2000. \textbf{A}.
    \item EM 386L -- \emph{Mathematical Methods in Applied Mechanics II}, Spring 2001. \textbf{A}.
    \item CS 395T -- \emph{High-Performance Graphics \& Visualization}, Spring 2002. \textbf{A}.
    \item CS 395T -- \emph{High-Performance \& Parallel Computing}, Spring 2002. \textbf{A}.    
    \end{enumerate}
\end{itemize}


\subsection*{PhD Exams}
\begin{enumerate}
\item ASE Written Qualifying Exams -- June 2001.
  \begin{itemize}
  \item Mathematics
  \item Fluid Dynamics
  \item Aerothermodynamics
  \end{itemize}
\item ASE Oral Qualifying Exam -- December, 2002.
%\item TICAM Area B (Numerical Analysis) Exam -- August 2002.
\end{enumerate}
    

\end{document}


  