\documentclass[12pt,pdftex]{article}

\topmargin -2.0cm
\oddsidemargin 0in
\evensidemargin 0in
%\footheight 0.0pt
\headheight 2\baselineskip
\textwidth 6.0in
\textheight 9.0in



\begin{document}
My research will focus on the development of novel finite element
simulation techniques for nonlinear transport applications such as
compressible and incompressible fluid dynamics and porous media
transport with reaction.  Specifically, the work will address issues
including:
\begin{enumerate}
  \item parallel computing and algorithm design for finite element
    applications with adaptive mesh refinement and dynamic load
    balancing,
  \item error estimation techniques for guiding adaptive mesh
    refinement,
  \item stabilization techniques for nonlinear, convection-dominated
    processes, and
  \item development of novel, high-order positivity preserving
    solution algorithms for reactive transport processes.
\end{enumerate}
Significant contributions of original research
are expected in each of these areas.  Application studies will include
both incompressible and compressible fluid dynamics applications and
porous media transport with reaction in the context of subsurface
contamination transport.  Specifically, physical phenomenon that
require adaptive grids such as shockwave/boundary layer interaction
will be considered. Algorithm performance and scalability on parallel
computing platforms and single node performance will also be examined.
Finally, the performance of new positivity preserving algorithms will
be examined theoretically, with numerical experiments, and with
comparisons to results in the literature.  The major consequence of
the proposed research will be efficient techniques for analyzing
nonlinear multiphysics problems with the finite element method,
particularly for fluid dynamics and porous media applications.
\end{document}